---
title:  "IVF Sample Analysis"
subtitle: "Day 3 vs Day 5 Embryo Transfer: "
date: "`r format(Sys.Date(), '%B %d, %Y')`"
author: ""
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: false
    theme: flatly
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 10,
  fig.height = 6,
  dpi = 150,
  out.width = "100%"
)
```

```{r libraries, include=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(scales)
library(patchwork)
library(ggridges)
library(geepack)
library(emmeans)
library(gtsummary)
library(gt)
library(openxlsx)
library(knitr)
library(kableExtra)
library(networkD3)
library(htmltools)

```

---

# Data Loading & Preparation

```{r load-data}
FILE_PATH <- "~/Library/CloudStorage/OneDrive-BowlingGreenStateUniversity/IVF Data/Boston_IVF.xlsx"

cycle_types <- c(
  "numeric","numeric","numeric","numeric","numeric",
  "numeric","numeric","text","numeric","numeric",
  "numeric","numeric","numeric","numeric","text",
  "text","numeric","numeric","numeric","numeric",
  "numeric","numeric","numeric"
)

cycles  <- read_excel(FILE_PATH, sheet = "cycles", col_types = cycle_types)

types <- c("text","numeric","text","numeric","numeric",
           "numeric","numeric","numeric","numeric","numeric",
           "numeric","numeric","text","text")

oocytes <- read_excel(FILE_PATH,
                      sheet   = "Oocyte‑level embryo development",
                      col_types = types)

best_codes  <- read_csv("~/Library/CloudStorage/OneDrive-BowlingGreenStateUniversity/IVF Data/sampled_patients177.csv")
sub_cycles  <- cycles  %>% dplyr::filter(Patient_Code %in% best_codes$patient_id)
sub_oocytes <- oocytes %>% dplyr::filter(Patient_Code %in% best_codes$patient_id)
```

```{r clean-data}
# ── cycles ────────────────────────────────────────────────────────────────
sub_cycles_clean <- sub_cycles %>%
  dplyr::mutate(
    Sperm_Origin       = as.factor(Sperm_Origin),
    Day_of_transfer    = as.factor(Day_of_transfer),
    Implantation       = as.factor(Implantation),
    `Pregnancy Status` = as.factor(`Pregnancy Status`),
    pregnancy    = factor(ifelse(`Pregnancy Status` == "pregnancy", 1, 0)),
    implantation = factor(ifelse(Implantation == "implantation", 1, 0))
  )

# ── oocytes ───────────────────────────────────────────────────────────────
blast_labels <- c(
  "1" = "Early Blastocyst",
  "2" = "Expanding Blastocyst",
  "3" = "Full Blastocyst",
  "4" = "Expanded Blastocyst",
  "5" = "Hatching Blastocyst",
  "6" = "Hatched Blastocyst"
)

sub_oocytes_clean <- sub_oocytes %>%
  dplyr::mutate(
    Day_Transfer        = as.factor(Day_Transfer),
    `Stage at Retrival` = as.factor(`Stage at Retrival`),
    `Fertility Status PN` = as.factor(`Fertility Status PN`),
    D5_Stage = suppressWarnings(as.numeric(D5_Stage)),
    D5_Stage = ifelse(D5_Stage < 0, NA, D5_Stage),
    D5_Stage = factor(D5_Stage, levels = 1:6, labels = unname(blast_labels))
  )

# ── analysis dataframe ────────────────────────────────────────────────────
df <- sub_cycles_clean %>%
  dplyr::mutate(
    Day = factor(Day_of_transfer, levels = c("3","5"),
                 labels = c("Day 3","Day 5")),
    Age_bin = cut(Patient_Age,
                  breaks = c(0, 30, 35, 40, 45),
                  labels = c("<30","30–34","35–39","40–44"),
                  right  = FALSE),
    BMI_bin = cut(Producer_BMI,
                  breaks = c(0, 18.5, 25, 30, 35, Inf),
                  labels = c("Underweight","Normal","Overweight",
                             "Obese I","Obese II+"),
                  right  = FALSE),
    MII_Rate = 100 * Sum_MII / N_Oocytes,
    Gravidity_cat = factor(ifelse(Gravidity == 0, "0", "1+"),
                           levels = c("0","1+")),
    Sperm_Origin   = droplevels(Sperm_Origin),
    Fert_Rate      = 100 * Fert_Rate,
    Cleavage_Rate  = 100 * Cleavage_Rate,
    Blast_Rate     = 100 * Blast_Rate
  )

df_gee <- df %>%
  dplyr::mutate(
    day_group = factor(Day_of_transfer, levels = c("3","5")),
    preg_bin  = as.integer(as.character(pregnancy)),
    impl_bin  = as.integer(as.character(implantation)),
    fh_bin    = ifelse(is.na(`FH>12wk`),    NA_integer_, as.integer(suppressWarnings(as.numeric(as.character(`FH>12wk`)))    >= 1)),
    birth_bin = ifelse(is.na(`#_Babies_(>24wks)`), NA_integer_, as.integer(suppressWarnings(as.numeric(as.character(`#_Babies_(>24wks)`))) >= 1)),
    sacs_bin  = ifelse(is.na(Sacs_Max),     NA_integer_, as.integer(suppressWarnings(as.numeric(as.character(Sacs_Max)))     >= 1)),
    gravidity_bin = as.integer(Gravidity > 0)
  ) %>%
  arrange(Patient_Code)

df_patient <- df %>%
  arrange(Patient_Code) %>%
  distinct(Patient_Code, .keep_all = TRUE)
```

```{r theme-colours, include=FALSE}
cols       <- c("Day 3" = "#fd5000", "Day 5" = "#00BFC4")
cols_group <- c("3"     = "#fd5000", "5"     = "#00BFC4")
col_labels <- c("3" = "Day 3", "5" = "Day 5")

theme_diss <- theme_classic(base_size = 12) +
  theme(
    plot.title         = element_text(size = 12, face = "bold", margin = margin(b=6)),
    plot.subtitle      = element_text(size = 10, color = "grey45", margin = margin(b=8)),
    plot.caption       = element_text(size = 9,  color = "grey55", hjust = 0),
    axis.title         = element_text(size = 11),
    axis.text          = element_text(size = 10),
    strip.text         = element_text(face = "bold", size = 10),
    strip.background   = element_blank(),
    legend.position    = "bottom",
    legend.title       = element_blank(),
    panel.grid.major.y = element_line(color = "grey92"),
    panel.grid.minor   = element_blank()
  )
```

```{r sem-helper, include=FALSE}
# SEM = SD / sqrt(n) — used in gtsummary statistic strings
sem <- function(x, na.rm = TRUE) {
  x <- x[!is.na(x)]
  sd(x) / sqrt(length(x))
}
```

```{r gee-functions, include=FALSE}
# ── GEE helper functions ──────────────────────────────────────────────────
gee_binary <- function(data, outcome, id_var = "Patient_Code", group = "day_group") {
  form  <- as.formula(paste(outcome, "~", group))
  model <- geeglm(form, id = data[[id_var]], data = data,
                  family = binomial(), corstr = "exchangeable")
  preds <- as.data.frame(
    emmeans(model, as.formula(paste("~", group)), type = "response")
  ) %>%
    rename_with(~ "estimate", any_of(c("prob", "rate", "emmean"))) %>%
    dplyr::rename(se = SE) %>%
    dplyr::mutate(lower = estimate - se, upper = estimate + se)
  pval  <- summary(model)$coefficients[2, "Pr(>|W|)"]
  stars <- ifelse(pval < 0.001,"***", ifelse(pval < 0.01,"**",
           ifelse(pval < 0.05,"*","")))
  preds %>%
    dplyr::mutate(outcome = outcome, model_type = "binary",
           p_value = pval, signif = stars) %>%
    dplyr::select(outcome, model_type, !!group, estimate, se, lower, upper, p_value, signif)
}

gee_continuous <- function(data, outcome, id_var = "Patient_Code", group = "day_group") {
  form  <- as.formula(paste(outcome, "~", group))
  model <- geeglm(form, id = data[[id_var]], data = data,
                  family = gaussian(), corstr = "exchangeable")
  preds <- as.data.frame(emmeans(model, as.formula(paste("~", group)))) %>%
    dplyr::rename(estimate = emmean, se = SE) %>%
    dplyr::mutate(lower = estimate - se, upper = estimate + se)
  pval  <- summary(model)$coefficients[2, "Pr(>|W|)"]
  stars <- ifelse(pval < 0.001,"***", ifelse(pval < 0.01,"**",
           ifelse(pval < 0.05,"*","")))
  preds %>%
    dplyr::mutate(outcome = outcome, model_type = "continuous",
           p_value = pval, signif = stars) %>%
    dplyr::select(outcome, model_type, !!group, estimate, se, lower, upper, p_value, signif)
}

make_gee_plot <- function(res, y_lab, title = NULL,
                          is_pct = FALSE, y_lim = NULL) {
  pval  <- unique(res$p_value)
  stars <- unique(res$signif)
  plab  <- if (is.na(pval)) "Day 5 only" else if (pval < 0.001) "p < 0.001" else paste0("p = ", round(pval, 3))
  annot <- if (is.na(pval)) plab else paste0("GEE Wald: ", plab, " ", stars)
  ymax  <- if (!is.null(y_lim)) y_lim else max(res$upper, na.rm = TRUE) * 1.20
  ymid  <- ymax * 0.97

  res %>%
    dplyr::mutate(day_group = factor(day_group, levels = c("3","5"))) %>%
    ggplot(aes(x = day_group, y = estimate,
               color = day_group, fill = day_group)) +
    geom_col(alpha = 0.35, width = 0.5) +
    geom_errorbar(aes(ymin = lower, ymax = upper),
                  width = 0.12, linewidth = 0.9) +
    geom_point(size = 2, shape = 21, stroke = 1.3) +
    annotate("text", x = 1.5, y = ymid,
             label = annot, size = 3.1, color = "grey30") +
    annotate("segment", x = 1, xend = 2,
             y = ymid * 0.92, yend = ymid * 0.92,
             color = "grey65", linewidth = 0.4) +
    scale_color_manual(values = cols_group) +
    scale_fill_manual(values  = cols_group) +
    scale_x_discrete(labels  = col_labels) +
    scale_y_continuous(
      labels = if (is_pct) scales::percent_format(accuracy = 1) else waiver(),
      limits = c(0, ymax),
      expand = expansion(mult = c(0, 0.02))
    ) +
    labs(x = NULL, y = y_lab,
         title = title) +
    theme_diss +
    theme(legend.position = "none")
}
```

---

# Sample Overview

```{r overview}
tibble(
  Metric = c("Total cycles","Unique patients",
             "Day 3 transfers","Day 5 transfers",
             "Overall pregnancy rate"),
  Value  = c(
    nrow(df),
    n_distinct(df$Patient_Code),
    sum(df$Day == "Day 3"),
    sum(df$Day == "Day 5"),
    paste0(round(mean(df$`Pregnancy Status` == "pregnancy") * 100, 1), "%")
  )
) %>%
  kable(align = c("l","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE)
```

---

# Demographics

> Patient-level analysis (deduplicated to one row per patient). Age and BMI are patient-level variables — Wilcoxon rank-sum tests are appropriate here.

## Summary table {.tabset}

### Table 1

```{r table1}
tbl_demo <- df_patient %>%
  dplyr::select(Day, Patient_Age, Producer_BMI, Gravidity_cat, Sperm_Origin) %>%
  tbl_summary(
    by = Day,
    label = list(
      Patient_Age   ~ "Age (years)",
      Producer_BMI  ~ "BMI (kg/m²)",
      Gravidity_cat ~ "Pregnancy history",
      Sperm_Origin  ~ "Sperm origin"
    ),
    statistic = list(
      all_continuous()  ~ "{mean} ± {sem}  [{median}, IQR {p25}–{p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 2),
    missing = "no"
  ) %>%
  add_p(test = list(all_continuous()  ~ "wilcox.test",
                    all_categorical() ~ "chisq.test"),
        pvalue_fun = ~ style_pvalue(.x, digits = 3)) %>%
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Table 1. Patient demographics by transfer day**
                  *Patient-level analysis (n = 1 row per patient)*") %>%
  modify_header(label ~ "**Variable**") %>%
  modify_footnote(p.value ~ "Wilcoxon rank-sum (continuous); chi-squared (categorical); patient-level")

tbl_demo
```

### Age bins

```{r table3}
tbl_age_day <- df_patient %>%
  count(Age_bin, Day) %>%
  group_by(Age_bin) %>%
  dplyr::mutate(Total = sum(n), Pct = round(100 * n / Total, 1)) %>%
  ungroup() %>%
  dplyr::mutate(cell = paste0(n, " (", Pct, "%)")) %>%
  dplyr::select(Age_bin, Day, cell, Total) %>%
  pivot_wider(names_from = Day, values_from = cell, values_fill = "0 (0%)") %>%
  dplyr::rename(`Age group` = Age_bin)

chi_age <- chisq.test(table(df_patient$Age_bin, df_patient$Day))

tbl_age_day %>%
  kable(caption = sprintf("Age group × transfer day. χ²=%.2f, p=%.3f%s",
                          chi_age$statistic, chi_age$p.value,
                          ifelse(chi_age$p.value<.001," ***",
                          ifelse(chi_age$p.value<.01," **",
                          ifelse(chi_age$p.value<.05," *"," (NS)"))))) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE) %>%
  footnote("n (% within age group); patient-level chi-squared test")
```

### BMI bins

```{r table4}
tbl_bmi_day <- df_patient %>%
  dplyr::filter(!is.na(BMI_bin)) %>%
  count(BMI_bin, Day) %>%
  group_by(BMI_bin) %>%
  dplyr::mutate(Total = sum(n), Pct = round(100 * n / Total, 1)) %>%
  ungroup() %>%
  dplyr::mutate(cell = paste0(n, " (", Pct, "%)")) %>%
  dplyr::select(BMI_bin, Day, cell, Total) %>%
  pivot_wider(names_from = Day, values_from = cell, values_fill = "0 (0%)") %>%
  dplyr::rename(`BMI category` = BMI_bin)

chi_bmi <- chisq.test(table(df_patient$BMI_bin, df_patient$Day))

tbl_bmi_day %>%
  kable(caption = sprintf("BMI category × transfer day. χ²=%.2f, p=%.3f%s",
                          chi_bmi$statistic, chi_bmi$p.value,
                          ifelse(chi_bmi$p.value<.001," ***",
                          ifelse(chi_bmi$p.value<.01," **",
                          ifelse(chi_bmi$p.value<.05," *"," (NS)"))))) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE) %>%
  footnote("n (% within BMI category); patient-level chi-squared test")
```

## Age plots

```{r plot-age, fig.height=5}
wt_age     <- wilcox.test(Patient_Age ~ Day, data = df_patient)
p_label_age <- sprintf("Wilcoxon p = %.4f%s", wt_age$p.value,
                        ifelse(wt_age$p.value<.001," ***",
                        ifelse(wt_age$p.value<.01," **",
                        ifelse(wt_age$p.value<.05," *"," (NS)"))))

p_viol_age <- ggplot(df_patient,
                     aes(x = Day, y = Patient_Age, fill = Day)) +
  geom_violin(alpha = 0.55, trim = FALSE, width = 0.8) +
  geom_boxplot(width = 0.18, alpha = 0.9, outlier.shape = NA, color = "grey25") +
  geom_jitter(aes(color = Day), width = 0.08, size = 1.3,
              alpha = 0.45, show.legend = FALSE) +
  stat_summary(fun = mean, geom = "point", shape = 23,
               size = 3, fill = "white", color = "grey25") +
  scale_fill_manual(values = cols) +
  scale_color_manual(values = cols) +
  annotate("text", x = 1.5, y = max(df_patient$Patient_Age, na.rm=TRUE) + 0.8,
           label = p_label_age, size = 3.5, color = "grey30") +
  labs(title = "Patient age by transfer day",
       subtitle = "Violin + boxplot; diamond = mean; patient-level Wilcoxon",
       x = NULL, y = "Age (years)", caption = "Whiskers = 1.5 × IQR") +
  theme_diss + theme(legend.position = "none")

p_ridge_age <- ggplot(df_patient,
                      aes(x = Patient_Age, y = Day, fill = Day)) +
  geom_density_ridges(alpha = 0.65, bandwidth = 1.4,
                      quantile_lines = TRUE, quantiles = c(0.25,0.5,0.75),
                      jittered_points = TRUE, point_shape = "|",
                      point_size = 2, point_alpha = 0.5,
                      position = position_raincloud(adjust_vlines = TRUE)) +
  scale_fill_manual(values = cols) +
  labs(title = "Age distribution", x = "Age (years)", y = NULL) +
  theme_diss + theme(legend.position = "none")

(p_viol_age | p_ridge_age) +
  plot_annotation(title = "Patient age by transfer day",
                  caption = "Patient-level analysis",
                  theme   = theme_diss)
```

## Age summary table

```{r tbl-age-summary}
df_patient %>%
  dplyr::group_by(Day) %>%
  dplyr::summarise(
    N       = n(),
    Mean    = round(mean(Patient_Age, na.rm=TRUE), 1),
    SE      = round(sd(Patient_Age, na.rm=TRUE)/sqrt(n()), 2),
    Median  = round(median(Patient_Age, na.rm=TRUE), 1),
    Q1      = round(quantile(Patient_Age, 0.25, na.rm=TRUE), 1),
    Q3      = round(quantile(Patient_Age, 0.75, na.rm=TRUE), 1),
    .groups = "drop"
  ) %>%
  dplyr::mutate(`Mean ± SE` = paste0(Mean, " ± ", SE),
                `Median [IQR]` = paste0(Median, " [", Q1, "–", Q3, "]")) %>%
  dplyr::select(Day, N, `Mean ± SE`, `Median [IQR]`) %>%
  kable(caption = "Age by transfer day (patient-level)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

## BMI plot

```{r plot-bmi, fig.height=5}
wt_bmi     <- wilcox.test(Producer_BMI ~ Day, data = df_patient)
p_label_bmi <- sprintf("Wilcoxon p = %.4f%s", wt_bmi$p.value,
                        ifelse(wt_bmi$p.value<.001," ***",
                        ifelse(wt_bmi$p.value<.01," **",
                        ifelse(wt_bmi$p.value<.05," *"," (NS)"))))

ggplot(df_patient, aes(x = Day, y = Producer_BMI, fill = Day)) +
  geom_violin(alpha = 0.55, trim = FALSE, width = 0.8) +
  geom_boxplot(width = 0.18, alpha = 0.9, outlier.shape = NA, color = "grey25") +
  geom_jitter(aes(color = Day), width = 0.08, size = 1.3,
              alpha = 0.35, show.legend = FALSE) +
  stat_summary(fun = mean, geom = "point", shape = 23,
               size = 3, fill = "white", color = "grey25") +
  geom_hline(yintercept = c(18.5,25,30,35),
             linetype = "dashed", color = "grey70", linewidth = 0.4) +
  annotate("text", x = 0.52, y = c(18.5,25,30,35) + 0.6,
           label = c("18.5","25","30","35"),
           size = 2.8, color = "grey55", hjust = 0) +
  scale_fill_manual(values = cols) +
  scale_color_manual(values = cols) +
  annotate("text", x = 1.5, y = max(df_patient$Producer_BMI, na.rm=TRUE) + 0.8,
           label = p_label_bmi, size = 3.5, color = "grey30") +
  labs(title = "BMI by transfer day",
       subtitle = "Dashed lines = WHO thresholds; patient-level Wilcoxon",
       x = NULL, y = "BMI (kg/m²)", caption = "Whiskers = 1.5 × IQR") +
  theme_diss + theme(legend.position = "none")
```


## BMI summary table

```{r tbl-bmi-summary}
df_patient %>%
  dplyr::filter(!is.na(Producer_BMI)) %>%
  dplyr::group_by(Day) %>%
  dplyr::summarise(
    N       = n(),
    Mean    = round(mean(Producer_BMI, na.rm=TRUE), 1),
    SE      = round(sd(Producer_BMI, na.rm=TRUE)/sqrt(n()), 2),
    Median  = round(median(Producer_BMI, na.rm=TRUE), 1),
    Q1      = round(quantile(Producer_BMI, 0.25, na.rm=TRUE), 1),
    Q3      = round(quantile(Producer_BMI, 0.75, na.rm=TRUE), 1),
    .groups = "drop"
  ) %>%
  dplyr::mutate(`Mean ± SE` = paste0(Mean, " ± ", SE),
                `Median [IQR]` = paste0(Median, " [", Q1, "–", Q3, "]")) %>%
  dplyr::select(Day, N, `Mean ± SE`, `Median [IQR]`) %>%
  kable(caption = "BMI by transfer day (patient-level)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

---

# Cycle-Level Outcomes

> Cycle-level outcomes use **GEE (generalised estimating equations)** with exchangeable correlation structure, clustered by `Patient_Code`, to account for patients contributing multiple cycles.

```{r run-gee, cache=TRUE}
# Day-5-only subset for blastocyst outcomes
df_gee_d5 <- dplyr::filter(df_gee, day_group == "5")

res_oocytes   <- gee_continuous(df_gee, "N_Oocytes")
res_mii       <- gee_continuous(df_gee, "Sum_MII")
res_miirate   <- gee_continuous(df_gee, "MII_Rate")
res_fert2pn   <- gee_continuous(df_gee, "N_Fert_2PN")
res_fert_rate <- gee_continuous(df_gee, "Fert_Rate")
res_nd3       <- gee_continuous(df_gee, "N_D3")
res_cleav     <- gee_continuous(df_gee, "Cleavage_Rate")
# N_D5 and Blast_Rate: Day-5-only descriptive summary (no Day 3 comparison)
summarise_d5 <- function(data, outcome) {
  x <- data[[outcome]]
  x <- x[!is.na(x)]
  tibble(
    outcome    = outcome,
    model_type = "d5_only",
    day_group  = "5",
    estimate   = mean(x),
    se         = sd(x) / sqrt(length(x)),
    lower      = mean(x) - sd(x) / sqrt(length(x)),
    upper      = mean(x) + sd(x) / sqrt(length(x)),
    p_value    = NA_real_,
    signif     = ""
  )
}
res_nd5   <- summarise_d5(df_gee_d5, "N_D5")
res_blast <- summarise_d5(df_gee_d5, "Blast_Rate")
res_transf    <- gee_continuous(df_gee, "Sum_Transferred")
res_preg      <- gee_binary(df_gee, "preg_bin")
res_impl      <- gee_binary(df_gee, "impl_bin")
res_sacs      <- gee_binary(df_gee, "sacs_bin")
res_birth     <- gee_binary(df_gee, "birth_bin")

# Implantation rate GEE (sacs/transferred, capped at 1)
impl_rate_gee_df <- df %>%
  dplyr::mutate(
    Sum_Transferred = as.numeric(Sum_Transferred),
    Sacs_Max        = as.numeric(Sacs_Max),
    impl_rate       = pmin(Sacs_Max / Sum_Transferred, 1)
  ) %>%
  dplyr::filter(!is.na(impl_rate), Sum_Transferred > 0) %>%
  dplyr::mutate(day_group = factor(Day_of_transfer, levels = c("3","5"))) %>%
  arrange(Patient_Code)
res_impl_rate <- gee_continuous(impl_rate_gee_df, "impl_rate")

label_map <- c(
  N_Oocytes       = "Oocytes retrieved",
  Sum_MII         = "MII oocytes",
  MII_Rate        = "MII rate",
  N_Fert_2PN      = "Fertilized (2PN)",
  Fert_Rate       = "Fertilization rate",
  N_D3            = "Embryos at Day 3",
  Cleavage_Rate   = "Cleavage rate",
  N_D5            = "Blastocysts at Day 5 (Day 5 cycles only)",
  Blast_Rate      = "Blastulation rate (Day 5 cycles only)",
  Sum_Transferred = "Embryos transferred",
  preg_bin        = "Clinical pregnancy",
  impl_bin        = "Implantation",
  impl_rate       = "Implantation rate (sacs/transferred)",
  sacs_bin        = "Gestational sacs ≥1",
  birth_bin       = "Live birth >24wk"
)

gee_results <- bind_rows(
  res_oocytes, res_mii, res_miirate, res_fert2pn, res_fert_rate,
  res_nd3, res_cleav, res_nd5, res_blast, res_transf,
  res_preg, res_impl, res_sacs, res_birth, res_impl_rate
) %>%
  dplyr::mutate(label = recode(outcome, !!!label_map))

gee_pvals <- gee_results %>% distinct(outcome, p_value, signif, label)
```

## Table 2 — Fertility outcomes

```{r table2}
tbl_fert <- df %>%
  dplyr::mutate(Day = factor(Day_of_transfer, levels = c("3","5"),
                      labels = c("Day 3","Day 5"))) %>%
  dplyr::select(Day, Sum_MII, N_Oocytes, MII_Rate, N_Fert_2PN,
         Fert_Rate, N_D3, Cleavage_Rate,
         Sum_Transferred, Implantation, `Pregnancy Status`,
         Sacs_Max, `#_Babies_(>24wks)`) %>%
  tbl_summary(
    by = Day,
    label = list(
      Sum_MII             ~ "MII oocytes (sum)",
      N_Oocytes           ~ "Oocytes retrieved",
      MII_Rate            ~ "MII rate (%)",
      N_Fert_2PN          ~ "Fertilized (2PN)",
      Fert_Rate           ~ "Fertilization rate (%)",
      N_D3                ~ "Embryos at Day 3",
      Cleavage_Rate       ~ "Cleavage rate (%)",
      Sum_Transferred     ~ "Embryos transferred",
      Implantation        ~ "Implantation",
      `Pregnancy Status`  ~ "Clinical pregnancy",
      Sacs_Max            ~ "Gestational sacs (max)",
      `#_Babies_(>24wks)` ~ "Live birth (>24 wks)"
    ),
    statistic = list(
      all_continuous()  ~ "{mean} ± {sem}  [{median}, IQR {p25}–{p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      all_continuous() ~ 1,
      c(MII_Rate, Fert_Rate, Cleavage_Rate) ~ 1
    ),
    missing = "no"
  ) %>%
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Table 2. Fertility and cycle outcomes by transfer day**
                  *p-values: GEE Wald test, exchangeable correlation, clustered by patient*") %>%
  modify_header(label ~ "**Variable**") %>%
  add_p(
    test = list(everything() ~ "wilcox.test"),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  ) %>%
  modify_table_body(
    ~ .x %>%
      dplyr::mutate(p.value = case_when(
        variable == "Sum_MII"           & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "Sum_MII"],
        variable == "N_Oocytes"         & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "N_Oocytes"],
        variable == "MII_Rate"          & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "MII_Rate"],
        variable == "N_Fert_2PN"        & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "N_Fert_2PN"],
        variable == "Fert_Rate"         & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "Fert_Rate"],
        variable == "N_D3"              & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "N_D3"],
        variable == "Cleavage_Rate"     & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "Cleavage_Rate"],
        variable == "Sum_Transferred"   & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "Sum_Transferred"],
        variable == "Implantation"      & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "impl_bin"],
        variable == "Pregnancy Status"  & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "preg_bin"],
        variable == "Sacs_Max"          & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "sacs_bin"],
        variable == "#_Babies_(>24wks)" & row_type == "label" ~
          gee_pvals$p_value[gee_pvals$outcome == "birth_bin"],
        TRUE ~ NA_real_
      ))
  ) %>%
  modify_footnote(
    p.value ~ "GEE Wald test; exchangeable correlation structure, clustered by Patient_Code"
  )

tbl_fert
```

## Figure 2 — Continuous outcomes (GEE)

```{r fig2-continuous, fig.height=13}
p1  <- make_gee_plot(res_oocytes,   "Oocytes retrieved",   title = "Oocytes retrieved")
p2  <- make_gee_plot(res_mii,       "MII oocytes",         title = "MII oocytes")
p10 <- make_gee_plot(res_miirate,   "MII rate (%)",        title = "MII rate",          is_pct = FALSE)
p3  <- make_gee_plot(res_fert2pn,   "Fertilized (2PN)",    title = "Fertilized 2PN")
p4  <- make_gee_plot(res_fert_rate, "Fertilization rate (%)", title = "Fertilization rate")
p5  <- make_gee_plot(res_nd3,       "Embryos at Day 3",    title = "Embryos at Day 3")
p6  <- make_gee_plot(res_cleav,     "Cleavage rate (%)",   title = "Cleavage rate")
p7  <- make_gee_plot(res_nd5,       "Blastocysts at Day 5",title = "Blastocysts at Day 5")
p8  <- make_gee_plot(res_blast,     "Blastulation rate (%)", title = "Blastulation rate")

(p1|p2|p10) /
  (p3|p4)/
  (p5|p6)+
  plot_annotation(
    title = "Oocyte retrival, Fertilization and Cleavage outcomes: Day 3 vs Day 5",
    subtitle = "GEE marginal means ± SE (exchangeable correlation, clustered by patient)",
    caption  = "*** p<0.001  ** p<0.01  * p<0.05",
    theme    = theme_diss
  )



```

## Continuous outcomes summary table

```{r tbl-continuous-summary}
cont_vars <- c("N_Oocytes","Sum_MII","MII_Rate","N_Fert_2PN",
               "Fert_Rate","N_D3","Cleavage_Rate","Sum_Transferred")
cont_labels <- c("Oocytes retrieved","MII oocytes","MII rate (%)","Fertilized (2PN)",
                 "Fertilization rate (%)","Embryos at Day 3","Cleavage rate (%)","Embryos transferred")

purrr::map2_dfr(cont_vars, cont_labels, function(v, lbl) {
  df %>%
    dplyr::group_by(Day) %>%
    dplyr::summarise(
      Mean   = round(mean(.data[[v]], na.rm=TRUE), 2),
      SE     = round(sd(.data[[v]], na.rm=TRUE)/sqrt(sum(!is.na(.data[[v]]))), 2),
      Median = round(median(.data[[v]], na.rm=TRUE), 2),
      Q1     = round(quantile(.data[[v]], 0.25, na.rm=TRUE), 2),
      Q3     = round(quantile(.data[[v]], 0.75, na.rm=TRUE), 2),
      .groups = "drop"
    ) %>%
    dplyr::mutate(Variable = lbl,
                  `Mean ± SE` = paste0(Mean, " ± ", SE),
                  `Median [IQR]` = paste0(Median, " [", Q1, "–", Q3, "]")) %>%
    dplyr::select(Variable, Day, `Mean ± SE`, `Median [IQR]`)
}) %>%
  pivot_wider(names_from = Day,
              values_from = c(`Mean ± SE`, `Median [IQR]`),
              names_glue = "{Day}: {.value}") %>%
  kable(caption = "Continuous outcomes by transfer day (cycle-level)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE)
```

## Figure 3 — Embryo transfer: Day 3 vs Day 5

```{r fig3-embryo-transfer, fig.height=6}
# ── descriptive stacked bar ────────────────────────────────────────────────
transf_counts <- df %>%
  dplyr::mutate(Transferred = factor(Sum_Transferred)) %>%
  count(Day, Transferred) %>%
  group_by(Day) %>%
  dplyr::mutate(pct = round(100 * n / sum(n), 1)) %>%
  ungroup()

p_transf_bar <- ggplot(transf_counts,
                       aes(x = Day, y = pct, fill = Transferred)) +
  geom_col(color = "white", linewidth = 0.3, alpha = 0.85) +
  geom_text(aes(label = ifelse(pct > 4, paste0(n,"\n(",pct,"%)"), "")),
            position = position_stack(vjust = 0.5),
            size = 3, color = "white", fontface = "bold") +
  # ggplot default discrete palette
  guides(fill = guide_legend(title = "N embryos\ntransferred")) +
  scale_y_continuous(labels = label_percent(scale = 1),
                     expand = expansion(mult = c(0, 0.03))) +
  labs(title    = "Embryos transferred — distribution",
       subtitle = "Stacked bars show proportion per transfer day",
       x = NULL, y = "% of cycles") +
  theme_diss + theme(legend.position = "right")

# ── GEE mean with CI ──────────────────────────────────────────────────────
p_transf_gee <- make_gee_plot(res_transf, "Mean embryos transferred",
                               title = "Embryos transferred (GEE)")

# ── combined ──────────────────────────────────────────────────────────────
(p_transf_bar | p_transf_gee) +
  plot_annotation(
    title = "Embryo transfer: Day 3 vs Day 5",
    subtitle = "Left: distribution of embryos transferred | Right: GEE marginal mean ± SE",
    caption  = "GEE Wald test; exchangeable correlation, clustered by patient",
    theme    = theme_diss
  )
```

## Figure 4 — Binary outcomes (GEE)

```{r fig4-binary, fig.height=6}
pb1 <- make_gee_plot(res_preg,  "Probability", title = "Clinical pregnancy",
                     is_pct = TRUE, y_lim = 1)
pb2 <- make_gee_plot(res_impl,  "Probability", title = "Implantation",
                     is_pct = TRUE, y_lim = 1)
pb3 <- make_gee_plot(res_sacs,  "Probability", title = "Gestational sacs ≥1",
                     is_pct = TRUE, y_lim = 1)
pb5 <- make_gee_plot(res_birth, "Probability", title = "Live birth >24wk",
                     is_pct = TRUE, y_lim = 1)

(pb1|pb2)/(pb3|pb5) +
  plot_annotation(
    title = "Binary outcomes: Day 3 vs Day 5",
    subtitle = "GEE estimated probabilities ± SE (exchangeable correlation, clustered by patient)",
    caption  = "*** p<0.001  ** p<0.01  * p<0.05",
    theme    = theme_diss
  )
```

## Binary outcomes summary table

```{r tbl-binary-summary}
bin_vars   <- c("preg_bin","impl_bin","sacs_bin","birth_bin")
bin_labels <- c("Clinical pregnancy","Implantation","Gestational sacs ≥1","Live birth >24wk")

purrr::map2_dfr(bin_vars, bin_labels, function(v, lbl) {
  df_gee %>%
    dplyr::group_by(day_group) %>%
    dplyr::summarise(
      N    = sum(!is.na(.data[[v]])),
      n    = sum(.data[[v]], na.rm=TRUE),
      Pct  = round(100 * mean(.data[[v]], na.rm=TRUE), 1),
      .groups = "drop"
    ) %>%
    dplyr::mutate(Variable = lbl,
                  Day = ifelse(day_group == "3", "Day 3", "Day 5"),
                  `n (%)` = paste0(n, " (", Pct, "%)")) %>%
    dplyr::select(Variable, Day, `n (%)`)
}) %>%
  pivot_wider(names_from = Day, values_from = `n (%)`) %>%
  kable(caption = "Binary outcomes by transfer day (cycle-level)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

## Table 3 — GEE results summary

```{r table-gee-results}
gee_results %>%
  dplyr::filter(day_group %in% c("3", "5")) %>%
  dplyr::mutate(
    `Mean ± SE` = paste0(round(estimate, 2), " ± ", round(se, 2)),
    Group       = ifelse(day_group == "3", "Day 3", "Day 5"),
    `p-value`   = ifelse(is.na(p_value), "—",
                    ifelse(p_value < 0.001, "<0.001", round(p_value, 3))),
    Sig         = ifelse(is.na(signif), "", signif)
  ) %>%
  dplyr::select(Outcome = label, Group, `Mean ± SE`, `p-value`, Sig) %>%
  pivot_wider(names_from = Group,
              values_from = `Mean ± SE`,
              names_prefix = "") %>%
  dplyr::rename(`p-value` = `p-value`, Significance = Sig) %>%
  # collapse duplicate p-value/sig columns from pivot
  dplyr::group_by(Outcome) %>%
  dplyr::summarise(
    `Day 3`        = dplyr::first(na.omit(`Day 3`)),
    `Day 5`        = dplyr::first(na.omit(`Day 5`)),
    `p-value`      = dplyr::first(na.omit(`p-value`)),
    Significance   = dplyr::first(na.omit(Significance)),
    .groups = "drop"
  ) %>%
  kable(caption = "**Table 3. GEE results: Mean ± SE by transfer day**",
        align = c("l","r","r","r","l")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE) %>%
  footnote("GEE Wald test; exchangeable correlation, clustered by Patient_Code.
            Rates reported as %. Implantation rate capped at 100%.")
```


---

# Oocyte-Level Analysis

```{r oocyte-function, include=FALSE}
summarise_factor_by_cycle <- function(data, var, id = "Cycle_Code") {
  var_sym <- sym(var); id_sym <- sym(id)
  counts <- data %>%
    dplyr::filter(!is.na(!!var_sym)) %>%
    group_by(!!id_sym, !!var_sym) %>%
    dplyr::summarise(n = n(), .groups = "drop")
  totals <- counts %>%
    group_by(!!id_sym) %>%
    dplyr::summarise(total = sum(n), .groups = "drop")
  lvls <- counts %>% distinct(!!var_sym) %>%
    pull(!!var_sym) %>% as.character() %>% sort()
  wide_n <- counts %>%
    pivot_wider(names_from = !!var_sym, values_from = n, values_fill = 0) %>%
    rename_with(~ paste0(.x,"_n"), all_of(lvls))
  wide_pct <- counts %>%
    left_join(totals, by = id) %>%
    dplyr::mutate(pct = round(100 * n / total, 1)) %>%
    dplyr::select(-n,-total) %>%
    pivot_wider(names_from = !!var_sym, values_from = pct, values_fill = 0) %>%
    rename_with(~ paste0(.x,"_pct"), all_of(lvls))
  paired <- as.vector(rbind(paste0(lvls,"_n"), paste0(lvls,"_pct")))
  wide_n %>%
    left_join(wide_pct, by = id) %>%
    left_join(totals, by = id) %>%
    arrange(!!id_sym) %>%
    relocate(!!id_sym, total) %>%
    dplyr::select(!!id_sym, total, all_of(paired))
}
```

```{r run-oocyte-tables}
tbl_stage_retrival <- summarise_factor_by_cycle(sub_oocytes_clean, "Stage at Retrival")
tbl_fertility_pn   <- summarise_factor_by_cycle(sub_oocytes_clean, "Fertility Status PN")
tbl_d5_stage       <- summarise_factor_by_cycle(sub_oocytes_clean, "D5_Stage")
tbl_d5_te          <- summarise_factor_by_cycle(sub_oocytes_clean, "D5_TE")
```

```{r oocyte-preview}
kable(head(tbl_stage_retrival, 10),
      caption = "Stage at retrieval — first 10 cycles shown") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                font_size = 11) %>%
  scroll_box(width = "100%")
```



```{r transfer-sacs-plot}
# ══════════════════════════════════════════════════════════════════════════
# PREP
# ══════════════════════════════════════════════════════════════════════════
transfer_sacs <- df %>%
  dplyr::mutate(
    Sum_Transferred = as.numeric(Sum_Transferred),
    Sacs_Max        = as.numeric(Sacs_Max)
  ) %>%
  dplyr::filter(!is.na(Sum_Transferred), !is.na(Sacs_Max)) %>%
  count(Day, Sum_Transferred, Sacs_Max) %>%
  group_by(Day, Sum_Transferred) %>%
  dplyr::mutate(
    pct_within_transferred = round(100 * n / sum(n), 1)
  ) %>%
  ungroup() %>%
  dplyr::mutate(
    Sacs_label = factor(Sacs_Max,
                        levels = sort(unique(Sacs_Max)),
                        labels = paste0(sort(unique(Sacs_Max)), " sac(s)"))
  )

# ══════════════════════════════════════════════════════════════════════════
# PLOT 1 — stacked bar: for each N transferred, how many sacs
# ══════════════════════════════════════════════════════════════════════════
p_stack <- ggplot(transfer_sacs,
                  aes(x = factor(Sum_Transferred),
                      y = pct_within_transferred,
                      fill = Sacs_label)) +
  geom_col(color = "white", linewidth = 0.3, alpha = 0.88) +
  geom_text(aes(label = ifelse(pct_within_transferred > 5,
                               paste0(n, "\n(", pct_within_transferred, "%)"), "")),
            position = position_stack(vjust = 0.5),
            size = 2.9, color = "white", fontface = "bold") +
  facet_wrap(~ Day, ncol = 2) +
  # ggplot default discrete palette
  guides(fill = guide_legend(title = "Gestational\nsacs")) +
  scale_y_continuous(labels = label_percent(scale = 1),
                     expand = expansion(mult = c(0, 0.03))) +
  labs(
    title    = "Gestational sacs by embryos transferred and transfer day",
    subtitle = "% within each 'embryos transferred' group",
    x        = "Embryos transferred (N)",
    y        = "% of cycles",
    caption  = "n shown inside bars where >5%"
  ) +
  theme_diss +
  theme(legend.position = "right")

# ══════════════════════════════════════════════════════════════════════════
# PLOT 2 — heatmap: transferred × sacs, split by day
# ══════════════════════════════════════════════════════════════════════════
p_heat <- ggplot(transfer_sacs,
                 aes(x = factor(Sum_Transferred),
                     y = factor(Sacs_Max),
                     fill = pct_within_transferred)) +
  geom_tile(color = "white", linewidth = 0.8) +
  geom_text(aes(label = paste0(n, "\n(", pct_within_transferred, "%)")),
            size = 3, color = "grey20") +
  facet_wrap(~ Day, ncol = 2) +
  scale_fill_gradient(low  = "#FFF5F0",
                      high = "#fd5000",
                      name = "% within\ntransferred group",
                      na.value = "grey95") +
  labs(
    title    = "Heatmap: embryos transferred vs gestational sacs",
    subtitle = "% within each 'N transferred' group per transfer day",
    x        = "Embryos transferred (N)",
    y        = "Gestational sacs (max)",
    caption  = "Cell = n cycles (% within row)"
  ) +
  theme_diss +
  theme(legend.position = "right",
        panel.grid      = element_blank())

# ══════════════════════════════════════════════════════════════════════════
# PLOT 3 — implantation rate: sacs / transferred per cycle
# ══════════════════════════════════════════════════════════════════════════
impl_rate <- df %>%
  dplyr::mutate(
    Sum_Transferred = as.numeric(Sum_Transferred),
    Sacs_Max        = as.numeric(Sacs_Max),
    impl_rate       = pmin(Sacs_Max / Sum_Transferred, 1)
  ) %>%
  dplyr::filter(!is.na(impl_rate), Sum_Transferred > 0)

# GEE for implantation rate (uses df_gee_impl created in run-gee chunk)

pval  <- unique(res_impl_rate$p_value)
stars <- unique(res_impl_rate$signif)
plab  <- sprintf("GEE Wald: p = %s %s",
                 ifelse(pval < .001, "<0.001", round(pval, 3)), stars)

p_impl_rate <- ggplot(impl_rate,
                      aes(x = Day, y = impl_rate, fill = Day)) +
 geom_violin(alpha = 0.45, trim = FALSE, width = 0.8) +
  geom_boxplot(width = 0.18, alpha = 0.85, outlier.shape = NA,
               color = "grey25") +
  geom_jitter(aes(color = Day), width = 0.08, size = 1.3,
              alpha = 0.35, show.legend = FALSE) +
  stat_summary(fun = mean, geom = "point", shape = 23,
               size = 3, fill = "white", color = "grey25") +
  scale_fill_manual(values  = cols) +
  scale_color_manual(values = cols) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  annotate("text", x = 1.5,
           y     = max(impl_rate$impl_rate, na.rm = TRUE) * 1.05,
           label = plab, size = 3.4, color = "grey30") +
  labs(
    title    = "Implantation rate per cycle",
    subtitle = "Sacs_Max / Sum_Transferred, capped at 100%; GEE Wald test",
    x        = NULL,
    y        = "Implantation rate (sacs / transferred)",
    caption  = "*** p<0.001  ** p<0.01  * p<0.05"
  ) +
  theme_diss + theme(legend.position = "none")

# ══════════════════════════════════════════════════════════════════════════
# COMBINED FIGURE
# ══════════════════════════════════════════════════════════════════════════
p_stack 

p_heat

  

 p_impl_rate
 
df %>%
  dplyr::mutate(
    Transferred = factor(as.numeric(Sum_Transferred)),
    Sacs        = factor(as.numeric(Sacs_Max))
  ) %>%
  dplyr::filter(!is.na(Transferred), !is.na(Sacs)) %>%
 count(Day, Transferred, Sacs) %>%
  group_by(Day, Transferred) %>%
  dplyr::mutate(pct = round(100 * n / sum(n), 1),
                cell = paste0(n, " (", pct, "%)")) %>%
  dplyr::select(-n, -pct) %>%
  pivot_wider(names_from = c(Day, Sacs), values_from = cell,
              values_fill = "0 (0%)", names_glue = "{Day} — {Sacs} sac(s)")%>%
  dplyr::rename(`Embryos transferred` = Transferred) %>%
  kable(
    caption = "Counts of embryos transferred × gestational sacs (max), by transfer day",
    align   = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  ) %>%
  add_header_above(
    # dynamically build the header: count how many sac levels exist per day
    c(" " = 1,
      "Day 3" = length(unique(df$Sacs_Max[df$Day == "Day 3" & !is.na(df$Sacs_Max)])),
      "Day 5" = length(unique(df$Sacs_Max[df$Day == "Day 5" & !is.na(df$Sacs_Max)]))
    )
  ) %>%
  footnote("Cells = number of cycles. Columns are max gestational sacs observed per cycle.")
```




## Implantation rate summary table

```{r tbl-impl-rate-summary}
impl_rate_gee_df %>%
  dplyr::group_by(day_group) %>%
  dplyr::summarise(
    N      = n(),
    Mean   = round(mean(impl_rate, na.rm=TRUE), 3),
    SE     = round(sd(impl_rate, na.rm=TRUE)/sqrt(n()), 3),
    Median = round(median(impl_rate, na.rm=TRUE), 3),
    Q1     = round(quantile(impl_rate, 0.25, na.rm=TRUE), 3),
    Q3     = round(quantile(impl_rate, 0.75, na.rm=TRUE), 3),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    Day           = ifelse(day_group == "3", "Day 3", "Day 5"),
    `Mean ± SE`   = paste0(round(Mean*100,1), "% ± ", round(SE*100,1), "%"),
    `Median [IQR]`= paste0(round(Median*100,1), "% [", round(Q1*100,1), "–", round(Q3*100,1), "%]")
  ) %>%
  dplyr::select(Day, N, `Mean ± SE`, `Median [IQR]`) %>%
  kable(caption = "Implantation rate by transfer day (sacs/transferred, capped at 100%)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```



# Export to Excel

```{r excel-export}
# Helper: make column names unique (case-insensitive) for openxlsx
dedup_colnames <- function(tbl) {
  nms <- colnames(tbl)
  # lowercase comparison to catch case-insensitive duplicates
  nms_lower <- tolower(nms)
  seen <- list()
  for (i in seq_along(nms_lower)) {
    key <- nms_lower[i]
    if (!is.null(seen[[key]])) {
      seen[[key]] <- seen[[key]] + 1L
      nms[i] <- paste0(nms[i], "_", seen[[key]])
    } else {
      seen[[key]] <- 1L
    }
  }
  colnames(tbl) <- nms
  tbl
}

export_tables_excel <- function(table_list, file) {
  wb <- createWorkbook()
  walk2(table_list, names(table_list), function(tbl, sheet_name) {
    sname <- substr(sheet_name, 1, 31)
    tbl   <- dedup_colnames(as.data.frame(tbl))   # deduplicate before writing
    addWorksheet(wb, sname)
    writeDataTable(wb, sname, tbl, tableStyle = "TableStyleMedium2")
    setColWidths(wb, sname, cols = 1:ncol(tbl), widths = "auto")
    freezePane(wb, sname, firstRow = TRUE)
    hs <- createStyle(fontColour = "#FFFFFF", fgFill = "#fd5000",
                      halign = "CENTER", textDecoration = "Bold",
                      border = "Bottom", borderColour = "#00BFC4")
    addStyle(wb, sname, hs, rows = 1, cols = 1:ncol(tbl), gridExpand = TRUE)
  })
  saveWorkbook(wb, file, overwrite = TRUE)
  cat("Saved:", file, "\n")
}

# ── output folder ─────────────────────────────────────────────────────────
out_dir <- file.path(
  "~/Library/CloudStorage/OneDrive-BowlingGreenStateUniversity/IVF Data",
  "IVF_Analysis_Output"
)
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

# ── flatten gtsummary tables ───────────────────────────────────────────────
tbl_demo_flat <- tbl_demo %>% as_tibble() %>% dplyr::select(-any_of("stat_0"))
tbl_fert_flat <- tbl_fert %>% as_tibble() %>% dplyr::select(-any_of("stat_0"))

# ── GEE summary ───────────────────────────────────────────────────────────
gee_summary <- gee_results %>%
  dplyr::mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  dplyr::select(label, model_type, day_group, estimate, se, lower, upper,
         p_value, signif)

export_tables_excel(
  list(
    # raw data
    "Data cycles"           = sub_cycles_clean %>%
                                dplyr::mutate(across(where(is.factor), as.character)),
    "Data oocytes"          = sub_oocytes_clean %>%
                                dplyr::mutate(across(where(is.factor), as.character)),
    # summary tables
    "Table1 Demographics"   = tbl_demo_flat,
    "Table2 Fertility"      = tbl_fert_flat,
    "Table3 Age Bin"        = tbl_age_day,
    "Table4 BMI Bin"        = tbl_bmi_day,
    # GEE results
    "GEE Results"           = gee_summary,
    # oocyte tables
    "Oocyte Stage Retrival" = tbl_stage_retrival,
    "Oocyte Fertility PN"   = tbl_fertility_pn,
    "Oocyte D5 Stage"       = tbl_d5_stage,
    "Oocyte D5 TE"          = tbl_d5_te
  ),
  file.path(out_dir, "All_Tables.xlsx")
)
```

---

# Statistical Methods Summary

```{r methods-table, echo=FALSE}
tibble(
  `Variable type`      = c("Age, BMI, gravidity","Age bin, BMI bin",
                            "Sperm origin","Cycle outcomes (continuous)",
                            "Pregnancy, implantation","Adjusted pregnancy"),
  `Unit of analysis`   = c("Patient","Patient","Patient",
                            "Cycle","Cycle","Cycle"),
  `Test used`          = c("Wilcoxon rank-sum","Chi-squared","Chi-squared",
                            "GEE (Gaussian)","GEE (binomial logit)",
                            "GEE multivariable"),
  `Reason`             = c("1 row/patient — independent","1 row/patient — independent",
                            "1 row/patient — independent",
                            "Clustered by patient","Clustered by patient",
                            "Clustered; age/BMI adjusted")
) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = TRUE)
```

> All GEE models used **exchangeable correlation structure** (assumes equal correlation between any two cycles from the same patient) and were clustered by `Patient_Code`. P-values are from **GEE Wald tests** (χ², 1 df).

---

*Report generated: `r format(Sys.time(), "%Y-%m-%d %H:%M")`*
